\documentclass[reqno, answers]{exam}

\usepackage{amssymb, amsmath, amsthm, stmaryrd}
\usepackage{fullpage, parskip}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{ulem}
\usepackage{hyperref}

\title{Introduction to Computational Math Spring 2026 \\ Homework 01}
\date{Due: Friday, January 30, 2025, 11:59 PM}
\author{}

\begin{document}
\maketitle
\thispagestyle{empty}

Textbook = Driscoll and Braun (\url{https://fncbook.com/})

\begin{enumerate}

\item Suppose you have a machine that works in \textit{base 10} instead of \textit{base 2}. In this case, the floating point representation of a non-zero real number would be of the form
\begin{align*}
  \pm(a + f) \times 10^n,
\end{align*}
where $f \in [0, 1)$ is the fractional part. This is nothing but the standard scientific notation. Note that this representation is unique for every non-zero real number.

\begin{enumerate}
  \item What values can the integer $a$ take?
  \item What values can each $b_i$ take in the fractional part $f = 0.b_1 b_2 b_3 \ldots$?
\end{enumerate}

Suppose the machine has precision $d = 3$, so that $f = 0.b_1 b_2 b_3$. Let $\mathbb{F}$ denote the set of all floating point numbers representable on this machine.

\begin{enumerate}[resume]
  \item What is the machine epsilon $\epsilon_{\text{mach}}$?
  \item How many elements of $\mathbb{F}$ are there in the real interval $[1, 100]$, including the endpoints? (Hint: Consider how many representable numbers exist for each exponent $n$.)
  \item What is the element of $\mathbb{F}$ closest to the real number $1/3$? 
  \item What is the smallest positive integer not in $\mathbb{F}$? (Hint: For what value of the exponent does the spacing between consecutive floating point numbers become larger than 1?)
\end{enumerate}

Suppose any information beyond the $d=3$ precision is simply lost (truncated rather than rounded). Let $a = 823.42$, $b = 723.45234$, and $c = 723.45$. Compute the following, showing intermediate steps:

\begin{enumerate}[resume]
  \item $a + (b - c)$
  \item $(a + b) - c$
\end{enumerate}

\begin{solution}
\begin{enumerate}
  \item Since $f \in [0,1)$, we require $a + f \in [1, 10)$ for a unique representation. Thus $a \in \{1, 2, 3, 4, 5, 6, 7, 8, 9\}$.
  
  \item Each $b_i$ is a base-10 digit, so $b_i \in \{0, 1, 2, 3, 4, 5, 6, 7, 8, 9\}$.
  
  \item Machine epsilon is the smallest $\epsilon$ such that $1 + \epsilon > 1$ in $\mathbb{F}$. The next representable number after $1.000 \times 10^0$ is $1.001 \times 10^0$. Hence $\epsilon_{\text{mach}} = 10^{-d} = 10^{-3} = 0.001$.
  
  \item The set $[1, 100]$ is the union of $[10^0, 10^1)$, $[10^1, 10^2)$, and the endpoint $100$. Each half-open interval $[10^n, 10^{n+1})$ contains $9 \times 10^d$ numbers, since the mantissa $a + f$ ranges over $[1, 10)$ with $10^d$ choices for each leading digit $a \in \{1, \ldots, 9\}$. Hence the count is $2 \cdot 9 \times 10^3 + 1 = 18001$.
  
\input{homework solutions for TAs/homework 01 solutions}  \item Since $1/3 \approx 0.333\ldots \in [10^{-1}, 10^0)$, take $n = -1$. Elements there have the form $10^{-1}(a + f)$ with $a \in \{1, \ldots, 9\}$ and $f = 0.b_1 b_2 b_3$. We need $a = 3$ and $f$ close to $0.333$. Candidates are $3.333 \times 10^{-1} = 0.3333$ and $3.334 \times 10^{-1} = 0.3334$. Since $1/3 = 0.333\ldots$, the nearer value is $3.333 \times 10^{-1}$.
  
  \item In $[10^n, 10^{n+1})$ the spacing between consecutive numbers is $10^{n-d}$. At $n = d = 3$, the spacing equals $10^0 = 1$, so all integers from $10^3$ to $10^4 - 1$ appear. At $n = d + 1 = 4$, the spacing is $10^1 = 10$, so only multiples of 10 appear. The smallest missing positive integer is therefore $10^{d+1} + 1 = 10001$.
  
\item Computing $a + (b - c)$:
  \begin{align*}
    b &= 723.45234 = 7.2345234 \times 10^{2} \quad \text{truncated:} \quad 7.234 \times 10^{2} \\
    c &= 723.45 = 7.2345 \times 10^{2} \quad \text{truncated:} \quad 7.234 \times 10^{2} \\
    b - c &= 7.234 \times 10^{2} - 7.234 \times 10^{2} = 0 \\
    a + (b - c) &= 823.4 + 0 = 8.234 \times 10^{2} \\
    \text{truncated:} & \quad \boxed{8.234 \times 10^{2}}
  \end{align*}
  
  \item Computing $(a + b) - c$:
  \begin{align*}
    a &= 823.4 = 8.234 \times 10^{2} \quad \text{truncated:} \quad 8.234 \times 10^{2} \\
    b &= 723.45234 = 7.2345234 \times 10^{2} \quad \text{truncated:} \quad 7.234 \times 10^{2} \\
    a + b &= 8.234 \times 10^{2} + 7.234 \times 10^{2} = 1546.8 = 1.5468 \times 10^3 \\
    \text{truncated:} & \quad 1.546 \times 10^3 \\
    (a + b) - c &= 1.546 \times 10^{3} - 7.234 \times 10^{2} = 822.6 = 8.226 \times 10^{2} \\
    \text{truncated:} & \quad \boxed{8.226 \times 10^{2}}
  \end{align*}

\end{enumerate}
\end{solution}


\item Textbook Exercise 1.1.1

\begin{solution}
  \begin{enumerate}
    \item The set $[1/2,4]$ is the union of $[2^{-1},1)$, $[1,2)$, $[2,4)$ and the endpoint $4$. Each half-open interval $[2^n,2^{n+1})$ contains $2^d$ numbers. Hence the count is $3\cdot 2^4 + 1 = 49$.
    \item Since $1/10 \in [2^{-4},2^{-3})$, take $n=-4$. Elements there have the form $2^{-4}(1+z/2^d)$ with $z\in\{0,\dots,2^d-1\}$. Now use $2^{-4}(1+z/16)$; taking $z\in\{9,10\}$ gives $25/256=0.09765625$ and $13/128=0.1015625$. The nearer to $0.1$ is $13/128$.
    \item In $[2^n,2^{n+1})$ the spacing is $2^{n-d}$. At $n=d$ the spacing equals $1$, so all integers from $2^d$ to $2^{d+1}-1$ appear. At $n=d+1$ the spacing is $2$ and only even integers appear, so the smallest missing positive integer is $2^{d+1}+1$. With $d=4$ this is $33$.
\end{enumerate}

\end{solution}

\item Textbook Exercise 1.1.4 (Express your answer using powers of 2. No need to write out all the digits.)
\begin{solution}
\begin{enumerate}
  \item The first single-precision number greater than 1 has the form $1 + 2^{-d} = 1 + 2^{-23}$. This also gives us $\epsilon_{\text{mach}} = 2^{-23}$.
  
  \item In $[2^n, 2^{n+1})$ the spacing between consecutive numbers is $2^{n-d} = 2^{n-23}$. At $n = d = 23$, the spacing equals $2^0 = 1$, so all integers from $2^{23}$ to $2^{24} - 1$ appear. At $n = d + 1 = 24$, the spacing is $2^1 = 2$, so only even integers appear. The smallest missing positive integer is therefore $2^{24} + 1$.
\end{enumerate}

\end{solution}

\item

Jupyter notebook.

\end{enumerate}



\end{document}